---
title: "Analysis of Teachers' Recruitment in Saudi Arabia Between 1437 and 1440"
subtitle: "Data from Saudi Open Data Office"
author: "areej mohamed othman"
date: "6 March 2021"
output: github_document
---

Group Number: team1

Group Members' Names (IDs):

1. aljawhara khalid(2201001960)   -- Team Leader
2asrar almutairi (2201000537) -- Member 1
3Ibtsam faleh.(2201002692) Elham khatim(2201001444)-- Member 2
4alaohoud faleh (2201002394)
5alkadi mutlaq (2201003964)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries 

Ensure you have the `tidyverse` and `readxl` packages installed by running the 2 lines of code below. Both the team leader and members should run the code below in the console of rstudio.

```{r libraries, echo=FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```


We then need to load the data from the MS Excel (.xlsx) file, named teachers_data.xlsx, to the R environment for processing.

*Note: Only Team Leader should run the code below*

```{r load_data, message = FALSE}
tdata <- read_excel("teachers_data.xlsx")
```

# Wherever you see the text '#Add some code', it means you should delete the comment and add your code for the task.

# 2 Marks will be deducted for every deadline missed.

## Task 1
\textcolor{red}{(8 points)}

```{r tidy_data}
tdata = rename(tdata,  administrative_regions = "المنطقة الإدارية")
tdata = rename(tdata,  school_level = "المرحلة")
tdata = rename(tdata,  authority = "السلطة")
tdata = rename(tdata,  Office_of_Education = "مكتب التربية")
tdata = rename(tdata,  sex = "الجنس")
tdata = rename(tdata,  School_type = "نوع المدرسة")
tdata = rename(tdata,  Type_of_education = "نوع التعليم")
tdata = rename(tdata,  school_system = "نظام الدراسة")
tdata = rename(tdata,  Saudi_teacher = "معلم سعودي")
tdata = rename(tdata,  Non_Saudi_teacher = "معلم غير سعودي")
tdata = rename(tdata,  Teachers = "معلمين جملة")
tdata = rename(tdata,  the_year = "السنة")
```

Team Leader should *Knit, commit, and push changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.*

\textcolor{red}{DEADLINE: END OF WEEK 9}

## Task2
\textcolor{red}{(8 points)}

Next Team Member should *Pull the changes made by Team Leader before proceeding.*

(a) What are the different administrative_regions (Al-Mandaqah Al-idariyyah) found in the data?

```{r admin_regions}
tdata %>% count(administrative_regions) %>% arrange(desc(n)) 
```

Comments about your answer:
the biggest regions ( alryad = 31617) and the smallest one ( alhodood alshamalya = 2072) . the most densely populated region is the one with the most employment of teachers (alryad)

(b) Based on your knowledge of the administrative regions in Saudi Arabia, is there any region that has not recruited any teacher during the period under review (1437 - 1440 Hijri)? List the Saudi regions shown in the data.

```{r not_recruited_admin_regions}
tdata %>% filter(the_year>=1437 ,the_year<=1440, Teachers==0)%>% count(administrative_regions)

tdata %>% filter(the_year>=1437 ,the_year<=1440, Non_Saudi_teacher==0)%>% count(administrative_regions)%>%arrange(desc(n)) 

tdata %>% filter(the_year>=1437 ,the_year<=1440, Saudi_teacher==0)%>% count(administrative_regions)%>%arrange(desc(n)) 
```

Comments about your answer:

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 10}

## Task3
\textcolor{red}{(20 points)}

The Next Team Member should *Pull the changes made by Member 1 before proceeding.*

(a) Saudi Arabian education is divided into different stages/levels (Al-Marhala), list the different levels showed in the data in descDEADLINE: ENDing order of frequency. (4 points)

```{r study_levels}
factor(tdata$levels_school)
table (tdata$levels_school)
count(tdata,  'levels_school')
tabl(tdata$levels_school,sort.group  =  "decreasing",cum.percent = TRUE)
tdata%>%arrange(desc(levels_school)) 

```

Comments about your answer: That we have 140438 (in) (n) of levels_school

(b) Which level recruit the most teachers? (4 points)


```{r most_teachers}
tdata%>%conunt(levels_school)%>%arrange(desc(n))
```

Comments about your answer:The largest level in recruitment is the primary Schools and the lowest level is the continue  learning so we should educate and qualify specialized teachers at the primary stage because it is The most important stage in terms of employment recruitment and also the most important in students lives

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

The Next Team member should *Pull the changes made by Team Leader before proceeding.*

(c) Plot the graph of region versus number of **Saudi** teachers recruited over the period. (4 points)

```{r plot_admin_regions_saudi}
ggplot(data=tdata,   aes(x = administrative_regions , y = saudi_teachers))
geom_point(stat=identity)

```

Comments about your answer:The most number of Saudi teachers has been employed in (riyad) and the fewest number of Saudi teachers have been employed in northern borders

(d) Plot the graph of region versus number of **Non-Saudi** teachers recruited over the period. (4 points)

```{r plot_admin_regions_non-saudi}
ggplot(data=tdata,   aes(x = administrative_regions , y = non_saudi_teachers))
geom_par(stat=identity)

```

Comments about your answer:The most number of  non Saudi teachers has been employed in (riyad) and the fewest number of Saudi teachers have been employed in Albahaa 

(e) Plot the graph of region versus number of **all teachers** recruited over the period. (4 points)

```{r plot_admin_regions_all}
ggplot(data=tdata,   aes(x = administrative_regions , y = total_saudi_teachers))
geom_point(stat=identity)

```

Comments about your answer:The most number of non or Saudi teachers has been employed in ( riyad ) and The fewest number of Saudi teathers have been employed in northern borders

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 12}

## Task 4
\textcolor{red}{(12 points)}

The Next Team member should *Pull the changes made previously before proceeding.*

(a) How many boys and how many girls schools are listed in the data? (4 points)

```{r no_boys_girls}
tdata%>%count(sex)
```

Comments about your answer:
the number of girls 'schools (77462) is more than the number of boys 'schools(62766)
(b) Between boys and girls schools, which recruit more teachers? (4 points)

```{r recruitment_boys_girls}
tdata%>%group_by(sex)%>%summarise_at(vars(Teachers),list(sum=sum))
```

Comments about your answer:
boys_schools_teachers=954028
girls_schools_teachers=1166895 


(c) What is the correlation between recruitment at boys schools and recruitment of girls schools? (4 points)

```{r plot_correlation_boys_girls}
ggplot(data=tdata, aes(x=sex, y= Teachers , color = sex)) +                              geom_bar(stat="identity")
```

Comments about your answer:
exceeds the teachers employed in boys' schools , so we must focus on training and teaching teachers at higher rates to prepare them for employment in girls 'schools.
the number of teachers employed in girls 'schools 

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 13}

## Task 5
\textcolor{red}{(12 points)}

The Next Team member should *Pull the changes made previously before proceeding.*

(a) How many teachers are recruited each year, 1437, 1438, 1439, and 1440? 

```{r teachers_per_year}
tdata%>%group_by(the_year)%>%summarise_at(vars(Teachers),list(sum=sum)) 
```

Comments about your answer:
the total of teachers (1437): (540883)
the total of teachers  (1438): (537965)
the total of teachers (1439): (525588)
the total of teachers (1440): (518726)

(b) Which region recruited the most teachers in each of the years under review? (4 points)

```{r teachers_per_region}
tdata%>%select(administrative_regions,Teachers,the_year)%>%filter(the_year=="1437")%>%group_by(administrative_regions)%>%summarise_at(vars(Teachers),list(sum=sum))%>%arrange(desc(sum))
tdata%>%select(administrative_regions,Teachers,the_year)%>%filter(the_year=="1438")%>%group_by(administrative_regions)%>%summarise_at(vars(Teachers),list(sum=sum))%>%arrange(desc(sum))
tdata%>%select(administrative_regions,Teachers,the_year)%>%filter(the_year=="1439")%>%group_by(administrative_regions)%>%summarise_at(vars(Teachers),list(sum=sum))%>%arrange(desc(sum))
tdata%>%select(administrative_regions,Teachers,the_year)%>%filter(the_year=="1440")%>%group_by(administrative_regions)%>%summarise_at(vars(Teachers),list(sum=sum))%>%arrange(desc(sum))
```

Comments about your answer:
the most region recruited in 1437 is   : Alriyad the recruited  : 130124
the most region recruited in 1438 is   : Alriyad the recruited  : 127911
the most region recruited in 1439 is   : Alriyad the recruited  : 123998
the most region recruited in 1440 is   : Alriyad the recruited  : 121030

(c) Plot the graph of number of teachers recruited vs the number of schools in an administrative region? What type or relationship exist between the two variables? (4 points)

```{r plot_teachers_administrative_regions}
#Add some code
```

Comments about your answer:

The Team Member should *Knit, commit, and push changes to GitHub with an appropriate commit message.*

\textcolor{red}{DEADLINE: END OF WEEK 14}

## If you are here, then congratulations!! You have completed the DSC 200 course project.
